---
title: "Histogrames"
author: "ETN"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r libraries, include = FALSE}
library(data.table)
library(tidyverse)
library(kableExtra)
library(scales)
load('/home/idiap/projects/zonamood/build_data.RData')
```

# Total

```{r total, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
data$resultat <- factor(data$resultat, levels = c('sense risc', "lleu", "moderat", 'greu'))
ggplot(data, aes(x = resultat, fill = resultat)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0(round((..count..)/sum(..count..) * 100, 1), "%")),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "risc",
       y = "Count") +
  theme_minimal()
```

```{r totalsuma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Create a histogram using hist()
hist(data$suma, breaks = 10, col = "steelblue", border = "white", xlab = "suma", ylab = "Frequency")

# Add a vertical red line
abline(v = 45, col = "red", lty = 2, lwd = 1)
```


# Benestar emocional

```{r benestar, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
benestar_emocional$resultat <- factor(benestar_emocional$resultat, levels = c('sense risc', "lleu", "moderat", 'greu'))
ggplot(benestar_emocional, aes(x = resultat, fill = resultat)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0(round((..count..)/sum(..count..) * 100, 1), "%")),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "risc",
       y = "Count") +
  theme_minimal()
```

```{r benestarsuma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Create a histogram using hist()
hist(benestar_emocional$suma, breaks = 10, col = "steelblue", border = "white", xlab = "suma", ylab = "Frequency")

# Add a vertical red line
abline(v = 12, col = "red", lty = 2, lwd = 1)
```

# Emocional

```{r emocional, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
emocional$resultat <- factor(emocional$resultat, levels = c('sense risc', "lleu", "moderat", 'greu'))
ggplot(emocional, aes(x = resultat, fill = resultat)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0(round((..count..)/sum(..count..) * 100, 1), "%")),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "risc",
       y = "Count") +
  theme_minimal()
```

```{r emocionalsuma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Create a histogram using hist()
hist(emocional$suma, breaks = 10, col = "steelblue", border = "white", xlab = "suma", ylab = "Frequency")

# Add a vertical red line
abline(v = 10, col = "red", lty = 2, lwd = 1)
```

# Social

```{r social, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
social$resultat <- factor(social$resultat, levels = c('sense risc', "lleu", "moderat", 'greu'))
ggplot(social, aes(x = resultat, fill = resultat)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0(round((..count..)/sum(..count..) * 100, 1), "%")),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "risc",
       y = "Count") +
  theme_minimal()
```

```{r socialsuma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Create a histogram using hist()
hist(social$suma, breaks = 10, col = "steelblue", border = "white", xlab = "suma", ylab = "Frequency")

# Add a vertical red line
abline(v = 6, col = "red", lty = 2, lwd = 1)
```

# Familiar

```{r familiar, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
familiar$resultat <- factor(familiar$resultat, levels = c('sense risc', "lleu", "moderat", 'greu'))
ggplot(familiar, aes(x = resultat, fill = resultat)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0(round((..count..)/sum(..count..) * 100, 1), "%")),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "risc",
       y = "Count") +
  theme_minimal()
```

```{r familiarsuma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Create a histogram using hist()
hist(familiar$suma, breaks = 10, col = "steelblue", border = "white", xlab = "suma", ylab = "Frequency")

# Add a vertical red line
abline(v = 3, col = "red", lty = 2, lwd = 1)
```

# Escolar

```{r escolar, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
relacional_escolar$resultat <- factor(relacional_escolar$resultat, levels = c('sense risc', "lleu", "moderat", 'greu'))
ggplot(relacional_escolar, aes(x = resultat, fill = resultat)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0(round((..count..)/sum(..count..) * 100, 1), "%")),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "risc",
       y = "Count") +
  theme_minimal()
```

```{r escolarsuma, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Create a histogram using hist()
hist(relacional_escolar$suma, breaks = 10, col = "steelblue", border = "white", xlab = "suma", ylab = "Frequency")

# Add a vertical red line
abline(v = 14, col = "red", lty = 2, lwd = 1)
```


# Total esferes + centinella


```{r esferes, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
data$resultat_tot = data$resultat
data = data %>% select(-suma, -resultat)
benestar_emocional$resultat_benestar = benestar_emocional$resultat
benestar_emocional = benestar_emocional %>% select(-suma, -resultat)
emocional$resultat_emocional = emocional$resultat
emocional = emocional %>% select(-suma, -resultat)
familiar$resultat_familiar = familiar$resultat
familiar = familiar %>% select(-suma, -resultat)
relacional_escolar$resultat_escolar = relacional_escolar$resultat
relacional_escolar = relacional_escolar %>% select(-suma, -resultat)
social$resultat_social = social$resultat
social = social %>% select(-suma, -resultat)

tables <- list(data, benestar_emocional, emocional, familiar, relacional_escolar, social)
esferes <- Reduce(left_join, tables)
esferes = esferes %>% mutate(risc = ifelse(resultat_tot!='sense risc' | resultat_benestar!='sense risc' | resultat_emocional!='sense risc' | resultat_familiar!='sense risc' | resultat_escolar!='sense risc' | resultat_social!='sense risc' | sol>=3 | social>=3 | no_grup_amics>=3 | casa_bien>=3 | actuar_davant_abus>=3, 'risc', 'no_risc'))
save(esferes, file='~/idiap/projects/zonamood/esferes.RData')
```

```{r histo_esferes, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
esferes$risc <- factor(esferes$risc, levels = c('risc', "no_risc"))
ggplot(esferes, aes(x = risc, fill = risc)) +
  geom_bar(stat = "count", position = "dodge") +
  geom_text(aes(label = paste0((..count..), '(',round((..count..)/sum(..count..) * 100, 1), "%", ')')),
            stat = "count", vjust = -0.5, color = "black") +
  scale_y_continuous(labels = comma, expand = c(0, 8)) +
  labs(title = "Distribution of Risks Total",
       x = "",
       y = "Count") +
  theme_minimal()
```

```{r histo_esferes_sexe, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
category_data <- esferes[, .(Count = .N), by = .(risc, SEXE)]
category_data[, Percentage := Count / sum(Count) * 100, by = risc]

ggplot(category_data, aes(x = risc, y = Count, fill = SEXE)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_text(aes(label = paste0(Count, '(', round(Percentage, 1), "%", ')')),
            position = position_dodge(width = 1), vjust = -0.5, color = "black") +
  labs(title = "Distribution of Risk by gender",
       x = "", y = "Count") +
  theme_minimal()

cont_table <- esferes %>%
  dplyr::count(SEXE, risc) %>%
  pivot_wider(names_from = risc, values_from = n, values_fill = 0)

# Perform the chi-square test
chi_square_result <- chisq.test(cont_table[, -1])
p_valor = chi_square_result$p.value
```

* No observem diferències significatives entre sexes, p-valor = `r p_valor`

```{r histo_esferes_coles, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
esferes <- esferes %>%
  mutate(cole = ifelse(str_starts(CODI, "GI"), "St.Gregori", "Calella"))

category_data <- esferes[, .(Count = .N), by = .(risc, cole)]
category_data[, Percentage := Count / sum(Count) * 100, by = risc]

ggplot(category_data, aes(x = risc, y = Count, fill = cole)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_text(aes(label = paste0(Count, '(', round(Percentage, 1), "%", ')')),
            position = position_dodge(width = 1), vjust = -0.5, color = "black") +
  labs(title = "Distribution of Risk by School",
       x = "", y = "Count") +
  theme_minimal()

cont_table <- esferes %>%
  dplyr::count(cole, risc) %>%
  pivot_wider(names_from = risc, values_from = n, values_fill = 0)

# Perform the chi-square test
chi_square_result <- chisq.test(cont_table[, -1])
p_valor = chi_square_result$p.value
```

* No observem diferències significatives entre col·legis, p-valor = `r p_valor`

# Gràfica preguntes

```{r histogrames_variables, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=10, fig.height=5}
numerical_vars <- names(data)[sapply(data, is.numeric)]
sums <- colSums(data[, ..numerical_vars])

# Create a data frame with variable names and sums
sum_df <- data.frame(variable = names(sums), sum = sums)

# Set the levels of the 'variable' column in the original order
sum_df$variable <- factor(sum_df$variable, levels = names(data))

# Create a new column for color groups
sum_df$esfera <- cut(
  as.numeric(sum_df$variable),
  breaks = c(0, 12, 24, 33, 39, Inf),
  labels = c("Benestar emocional", "emocional", "social", "familiar", "escolar")
)

# Create a bar plot with different colors for each group
plot <- ggplot(sum_df, aes(x = variable, y = sum, fill = esfera)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Sum of Variables", x = "Variables", y = "Sum") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_fill_manual(
    values = c("Benestar emocional" = "steelblue", "emocional" = "darkorange",
               "social" = "green", "familiar" = "purple", "escolar" = "red")
  )

# Display the plot
plot

```

## Per sexe

```{r hostogrames_variables_genere, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=10, fig.height=5}
males = data[SEXE=='Masculí']
females = data[SEXE=='Femení']

# Plot the first histogram with normal distribution
numerical_vars <- names(males)[sapply(males, is.numeric)]
sums <- colSums(males[, ..numerical_vars])
sum_df <- data.frame(variable = names(sums), sum = sums)
sum_df$variable <- factor(sum_df$variable, levels = names(data))
sum_df$esfera <- cut(
  as.numeric(sum_df$variable),
  breaks = c(0, 12, 24, 33, 39, Inf),
  labels = c("Benestar emocional", "emocional", "social", "familiar", "escolar")
)
ggplot(sum_df, aes(x = variable, y = sum, fill = esfera)) +
  geom_bar(stat = "identity") +
  labs(title = "Homes", x = "Variables", y = "Sum") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_fill_manual(
    values = c("Benestar emocional" = "steelblue", "emocional" = "darkorange",
               "social" = "green", "familiar" = "purple", "escolar" = "red")
  )+
  ylim(0, 80)

# Plot the second histogram with normal distribution
numerical_vars <- names(females)[sapply(females, is.numeric)]
sums <- colSums(females[, ..numerical_vars])
sum_df <- data.frame(variable = names(sums), sum = sums)
sum_df$variable <- factor(sum_df$variable, levels = names(data))
sum_df$esfera <- cut(
  as.numeric(sum_df$variable),
  breaks = c(0, 12, 24, 33, 39, Inf),
  labels = c("Benestar emocional", "emocional", "social", "familiar", "escolar")
)
ggplot(sum_df, aes(x = variable, y = sum, fill = esfera)) +
  geom_bar(stat = "identity") +
  labs(title = "Dones", x = "Variables", y = "Sum") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_fill_manual(
    values = c("Benestar emocional" = "steelblue", "emocional" = "darkorange",
               "social" = "green", "familiar" = "purple", "escolar" = "red")
  )+
  ylim(0, 80)

```

### Piramide

```{r piramide_poblacio_sexe, results='asis', warning=FALSE}
taula = data %>% select(-CODI, -resultat_tot)
taula = taula[SEXE!='Altres']
# dades = taula %>% select(-cole)
melted_dades <- melt(taula, id.vars = "SEXE")
result <- melted_dades %>% group_by(variable, SEXE) %>% summarize_if(is.numeric, sum, na.rm=TRUE)
setnames(result, c("variable", "sexe", "Sum"))


# Create a basic bar chart for one gender
basic_plot <-  ggplot(
    result, 
    aes(
        x = variable, 
        fill = sexe, 
        y = ifelse(
            test = sexe == "Masculí", 
            yes = -Sum, 
            no = Sum
            )
        )
    ) + 
geom_bar(stat = "identity") 

# Create population pyramids for both genders and combine them
population_pyramid <- basic_plot +
  scale_y_continuous(
    labels = abs, 
    limits = max(result$Sum) * c(-1,1)
  ) + 
  coord_flip() + 
  theme_minimal() +
  scale_fill_manual(values = c('Masculí' = "lightblue", 'Femení' = "lightcoral"))+
  labs(
    x = "variable", 
    y = "Sum", 
    fill = "sexe", 
    title = "Population Pyramid"
  )
population_pyramid
```

* Les següents variables presenten diferències significatives entre sexes segons chi-squared test:

```{r chi-test_sexe}
count_table <- data %>% select(-CODI, -resultat_tot)
count_table = count_table[SEXE!='Altres']
# Get the names of the categorical variables (excluding "SEXE")
cat_vars <- names(count_table)[-1]

# Perform chi-squared tests for each variable
results <- lapply(cat_vars, function(var) {
  contingency_table <- table(count_table$SEXE, count_table[[var]])
  chisq.test(contingency_table)
})
# Extract the p-values from the chi-squared tests
p_values <- sapply(results, function(res) res$p.value)

# Identify statistically significant differences
significant_vars <- cat_vars[p_values < 0.05]

# Print the significant variables
print(significant_vars)


```

## Per col·legi

```{r hostogrames_variables_coles, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=10, fig.height=5}
data = esferes %>% select(-CODI, -SEXE, -resultat_tot, -resultat_benestar, -resultat_emocional, -resultat_familiar, -resultat_escolar, -resultat_social, -risc)
calella = data[cole=='Calella']
gregori = esferes[cole=='St.Gregori']

# Plot the first histogram with normal distribution
numerical_vars <- names(calella)[sapply(calella, is.numeric)]
sums <- colSums(calella[, ..numerical_vars])
sum_df <- data.frame(variable = names(sums), sum = sums)
sum_df$variable <- factor(sum_df$variable, levels = names(data))
sum_df$esfera <- cut(
  as.numeric(sum_df$variable),
  breaks = c(0, 10, 22, 31, 37, Inf),
  labels = c("Benestar emocional", "emocional", "social", "familiar", "escolar")
)
ggplot(sum_df, aes(x = variable, y = sum, fill = esfera)) +
  geom_bar(stat = "identity") +
  labs(title = "Calella", x = "Variables", y = "Sum") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_fill_manual(
    values = c("Benestar emocional" = "steelblue", "emocional" = "darkorange",
               "social" = "green", "familiar" = "purple", "escolar" = "red")
  )+
  ylim(0, 80)

# Plot the second histogram with normal distribution
numerical_vars <- names(gregori)[sapply(gregori, is.numeric)]
sums <- colSums(gregori[, ..numerical_vars])
sum_df <- data.frame(variable = names(sums), sum = sums)
sum_df$variable <- factor(sum_df$variable, levels = names(data))
sum_df$esfera <- cut(
  as.numeric(sum_df$variable),
  breaks = c(0, 10, 22, 31, 37, Inf),
  labels = c("Benestar emocional", "emocional", "social", "familiar", "escolar")
)
ggplot(sum_df, aes(x = variable, y = sum, fill = esfera)) +
  geom_bar(stat = "identity") +
  labs(title = "St.Gregori", x = "Variables", y = "Sum") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  scale_fill_manual(
    values = c("Benestar emocional" = "steelblue", "emocional" = "darkorange",
               "social" = "green", "familiar" = "purple", "escolar" = "red")
  )+
  ylim(0, 80)

```

### Piramide

```{r piramide_poblacio_colegis, results='asis', warning=FALSE}
taula = data
# dades = taula %>% select(-cole)
melted_dades <- melt(taula, id.vars = "cole")
result <- melted_dades %>% group_by(variable, cole) %>% summarize_if(is.numeric, sum, na.rm=TRUE)
setnames(result, c("variable", "cole", "Sum"))


# Create a basic bar chart for one gender
basic_plot <-  ggplot(
    result, 
    aes(
        x = variable, 
        fill = cole, 
        y = ifelse(
            test = cole == "Calella", 
            yes = -Sum, 
            no = Sum
            )
        )
    ) + 
geom_bar(stat = "identity") 

# Create population pyramids for both genders and combine them
population_pyramid <- basic_plot +
  scale_y_continuous(
    labels = abs, 
    limits = max(result$Sum) * c(-1,1)
  ) + 
  coord_flip() + 
  theme_minimal() +
  scale_fill_manual(values = c('Calella' = "lightblue", 'St.Gregori' = "lightcoral"))+
  labs(
    x = "variable", 
    y = "Sum", 
    fill = "cole", 
    title = "Population Pyramid"
  )
population_pyramid
```

* Les següents variables presenten diferències significatives entre col·legis segons chi-squared test:

```{r chi-test_cole}
count_table <- data 
# Get the names of the categorical variables (excluding "SEXE")
cat_vars <- names(count_table)[-47]

# Perform chi-squared tests for each variable
results <- lapply(cat_vars, function(var) {
  contingency_table <- table(count_table$cole, count_table[[var]])
  chisq.test(contingency_table)
})
# Extract the p-values from the chi-squared tests
p_values <- sapply(results, function(res) res$p.value)

# Identify statistically significant differences
significant_vars <- cat_vars[p_values < 0.05]

# Print the significant variables
print(significant_vars)


```